---
title: "ESM 232 Modeling - Rabbits and Hawks"
subtitle: "Population Matrix Assignment"
author: "Shuhan Song, Ruoyu Wang"
date: "5/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = "center")
library(tidyverse)
library(sensitivity)
library(patchwork)
```

### 1. Rabbit population in 20 years with initial parameter setting

```{r initial.condition}
# set parameters
fert_rab = c(0,2,6,1) # fertility
surv_rab = c(0.8,0.85,0.65,0.1) #survivability
ini = c(0,0,10,0) # initial population
nyears = 20
```

```{r evolve_pop}
# call the function
source("R/evolve_pop.R")

# apply the function with initial setting
rab_pop=evolve_pop(fert_rab, surv_rab, ini, nyears)

# Total population of rabbits after 20 years
rab_pop1 <- rab_pop$poptot[20]
rab_pop1

# Population of young rabbits after 20 years
rab_pop$popbyage[1,20]
```

### 2. Sensitivity analysis about rabbit population with young and sub-adult survivability variations due to hawks

#### a. Sobel samples and wrapper function

```{r sobel.df}
# initialize two samples of parameters
nsample = 200

sample1 = data.frame(p01 = runif(n=nsample,
                                 min=0.65,
                                 max=0.75),
                     p12 = runif(n=nsample,
                                 min=0.75,
                                 max=0.8))

sample2 = data.frame(p01 = runif(n=nsample, 
                                 min=0.65, 
                                 max=0.75),
                     p12 = runif(n=nsample,
                                 min=0.75,
                                 max=0.8))

# generate sobel samples
sens_rabsur=soboljansen(model = NULL, sample1, sample2, nboot = 100)

# check the sobel samples we created
head(sens_rabsur$X)
nsim=nrow(sens_rabsur$X)
```

```{r constants}
# constant inputs

## survivability
p23 = 0.65
p34 = 0.1

## fertility rate
f1 = 0.0
f2 = 2
f3 = 6
f4 = 1

## initial population
ini = c(0,0,10,0)

## nstep
nyears = 20
```

```{r wrapper.apply}
# create wrapper function
p_wrapper = function(p01, p12, p23, p34,
                     f1, f2, f3, f4, 
                     use_func, initialpop, nstep){
  
  fertility=c(f1,f2,f3,f4)
  
  survivability= c(p01,p12,p23,p34)
  
  res = use_func(survivability =survivability,
                 fertility = fertility,
                 initialpop=initialpop, 
                 nstep=nstep)
  
# now return the final population total
return(finalpop=res$poptot[nstep])
}

# apply the wrapper function with sobel samples and constant parameters
res = sens_rabsur$X %>%
  pmap_dbl(p_wrapper, 
           p23=p23, p34=p34,
           f1=f1, f2=f2,
           f3=f3, f4=f4,
           initialpop=ini, 
           nstep=nyears, 
           use_func=evolve_pop)

# store results in a dataframe 
res_df <- as.data.frame(res) %>% 
  rename(finalpop=res)
```

#### b. Rabbit final population under hawks' impact versus the original final population

```{r boxplot, fig.width=4, fig.height=4}
# boxplot for final pop
ggplot(data.frame(finalpop=res), 
       aes(x="", y=finalpop) )+
  geom_boxplot(fill = "skyblue")+
  geom_hline(yintercept = rab_pop1,
             color = "darkblue",
             size = 1,
             linetype = "longdash")+
  geom_text(aes(y = rab_pop1-300000, 
                label = sprintf("Original Final Pop %4.0f",rab_pop1)),
            color = "darkblue")+
  theme_bw()+
  theme(axis.title.x = element_blank())+
  labs(y="Final Rabbit Population",
       title = "Boxplot of Final Rabbit Population")
```

#### c. Find out which parameter has more influence on rabbit population

```{r ses}
# give our results to sensitivity structure
sens_rabsur=tell(sens_rabsur, res)

# look at results
sens_rabsur$S
sens_rabsur$T
```

Both the main effect and total effect results show that the survivability of Young age group (p01) has significantly higher impact on the final population than the Sub-adult age group (p12).

The following scatterplots of parameters versus population outputs also support the finding.

```{r sens.plot, fig.width=11, fig.height=4}
# graph the most sensitive parameter
tmp = cbind.data.frame(sens_rabsur$X, pop12=sens_rabsur$y)

p12 <- ggplot(tmp, aes(p12, pop12))+
  geom_point()+
  labs(x="Survivability of p12",
       y="Population after 20 years")+
  theme_bw()

p01 <- ggplot(tmp, aes(p01, pop12))+
  geom_point()+
  labs(x="Survivability of p01",
       y="Population after 20 years")+
  theme_bw()

p01 + p12
```

